---
title: "Explorations"
author: "Alejandra Garcia Isaza"
date: '2022-10-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)

theme_set(theme_minimal())
```

# Loading the dataset
```{r}
d <- read_sav(here("nopublish", "PARQ.Stacked.sav"))
```
# Initial cleaning & filtering

"Participants were 241 families, each with at least one youth
in grade 5, 6, 7, or 8. Since the study included both single-
and two-parent families, outcome analyses reported
here focus on data from the primary caregiver in the family
as reported by parents (i.e., 237 mothers and 4 fathers)"

2 = dad
3 = mom

# these are the four dads
LY566
LY807
LY813
LY858

```{r}

# d2 <- d %>%
#   filter(wave == 1 | wave == 2)

d2 <- d %>%
  filter(wave == 2) # 424 participants total
```

# selecting only primary caregivers
```{r}

length(unique(d2$FAMILY)) # 241 primary caregiver per family

data.frame(table(d2$FAMILY))

only_moms <- d2 %>%
  filter(resp == 3) # 237 were considered primary moms

only_dads <- d2 %>%
  filter(FAMILY == "LY566" | FAMILY == "LY807" | FAMILY == "LY813" | FAMILY == "LY858") # 4 single dads

d3 <- rbind(only_moms, only_dads) %>%
  arrange(FAMILY) # this is the data set to work with
```
```{r}
d3 %>%
  filter(MISSPARQ == 1) # family LY564 didn't complete this questionnaire... Remove from analyses?
```


--------------------------
## EFA decisions 
--------------------------

"The tetrachoric correlation is a specific type of polychoric correlation that obtains
when both observed variables are dichotomous" (Flora et al., 2012)

# HW involvement
```{r}
hw <- d3 %>%
  select(PVKA16.1, PVKA16.2, PVKA16.3, PVKA16.4, PVKA16.5, PVKA16.6, PVKA16.7, PVKA16.8, PVKA16.9, PVKA16.10, PVKA16.11, PVKA16.12, PVKA16.13, PVKA16.14, PVKA16.15, PVKA16.16)
```
0 = Not apply
1 = apply

* low loading
** very low loading


**Youth has full responsibility for getting homework home; try to stay out of it (PVKA16.1)
**I have given up (helping youth get homework done)
Youth has certain time to do homework
*Youth has certain place to do homework
I sit with youth while s|he does homework
I help youth when s|he is stuck on homework
I help youth by quizzing her|him to prepare for tests, etc.
I give youth incentives, like special privileges, special treats, etc.
I use sanctions, punishment, discipline actions
*I remind youth to do homework
I check with youth to be sure homework is done
I check, correct homework with youth
I check with teacher about how well homework is done
I help access other resources (library, art supplies, computer labs etc.)
I help youth practice skills (read, go over problems, etc.)
I limit noise or distractions when youth is doing homework (PVKA16.16)


```{r}
fa(hw, n.obs = NA, rotate = "none", fm = "uls", cor = "tet") # here still need to investigate if uls is the best extraction method for tetrachocric cor
```
```{r}
# PARALLEL ANALYSIS

fa.parallel(hw, n.obs=NULL, fm="uls", fa="fa", nfactors=1, main="Parallel Analysis Scree Plots", n.iter=20, error.bars=FALSE, se.bars=FALSE, SMC=FALSE, ylabel=NULL, show.legend=TRUE, sim=TRUE, quant=.95, cor="tet", use="pairwise", plot=TRUE, correct=.5) # suggested number of factors 4 (with our without first 2 items)
```



# Homework Involvement (17) -- JUNTOS as model
```{r eval=FALSE}
# # Homework Involvement (17)
# scale_7 <- efa_vars %>%
#   select(q82_p1:q98_p1)
# 
# # Poly corr matrix
# poly_scale_7 <- polychoric(scale_7)
# poly_scale_7_mat <- data.frame(poly_scale_7$rho)
# 
# # scree plot
# scree(poly_scale_7_mat,factors=TRUE,pc=FALSE,main="Scree plot",hline=NULL,add=FALSE)
# 
# # EFA
# fa(scale_7, n.obs = NA, rotate = "none", fm = "uls", cor = "poly")
```

```{r eval=FALSE}
#removing 82, 83, 92
# scale_7.1 <- efa_vars %>%
#   select(q84_p1:q91_p1, q93_p1:q98_p1)
# 
# # PARALLEL ANALYSIS - Homework Involvement (17)
# fa.parallel(scale_7.1, n.obs=NULL, fm="uls", fa="fa", nfactors=1, main="Parallel Analysis Scree Plots", n.iter=20, error.bars=FALSE, se.bars=FALSE, SMC=FALSE, ylabel=NULL, show.legend=TRUE, sim=TRUE, quant=.95, cor="poly", use="pairwise", plot=TRUE, correct=.5) # parrallel analysis suggest 4 factors
# 
# # 4 factors
# fa(scale_7.1, nfactors = 4, n.obs = NA, rotate = "quartimin", fm = "uls", cor = "poly")
# 
# # 3 factors
# fa(scale_7.1, nfactors = 3, n.obs = NA, rotate = "quartimin", fm = "uls", cor = "poly")
```

```{r eval=FALSE}
#Reliability: Homework Involvement (17)
# scale_7.1_fa_1 <- efa_vars %>%
#   select(q86_p1:q89_p1, q94_p1, q96_p1, q97_p1)
# 
# alpha(scale_7.1_fa_1) # 0.78
# 
# scale_7.1_fa_2 <- efa_vars %>%
#   select(q84_p1, q85_p1, q93_p1, q95_p1)
# 
# alpha(scale_7.1_fa_2) # 0.70 (0.77 if removing 95)
# 
# scale_7.1_fa_3 <- efa_vars %>%
#   select(q90_p1, q91_p1, q98_p1) 
# 
# alpha(scale_7.1_fa_3) # 0.59
```

